---
title: "scpGUI"
author:
    - name: Léopold Guyot
output:
    BiocStyle::html_document:
        self_contained: yes
        toc: true
        toc_float: true
        toc_depth: 2
        code_folding: show
bibliography: scpGUI.bib
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('scpGUI')`"
vignette: >
  %\VignetteIndexEntry{scpGUI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# scpGUI

`scpGUI` is a shiny application that serves as a graphical interface for the [`scp` package](https://bioconductor.org/packages/release/bioc/html/scp.html) (@scpVanderaa). It leverages the power of Shiny and ShinydashboardPlus to provide a user-friendly interface for SCP data analysis. The app is organized into multiple sections, each dedicated to a specific aspect of SCP data processing. These sections are documented in separate vignettes to make it easy for users to navigate and find information relevant to their needs.

# The `scp` Package

`scpGUI` is a graphical interface for the [`scp` package](https://bioconductor.org/packages/release/bioc/html/scp.html) it is therefore recommended to understand how `scp` will handle the SCP data. In this short section, the general idea behind `scp` is presented refer to the [documentation](https://uclouvain-cbio.github.io/scp/index.html) of the package for more information.

The [`scp` package](https://bioconductor.org/packages/release/bioc/html/scp.html) is a robust tool for processing and analyzing MS-based SCP data. It operates on a specialized data structure that wraps [`QFeatures`](https://rformassspectrometry.github.io/QFeatures/) objects (@Gatto2020-ry) around [`SingleCellExperiment`](http://bioconductor.org/packages/release/bioc/html/SingleCellExperiment.html) objects (@Amezquita2019-bf). This unique data structure can be conceptualized as Matryoshka dolls, where SingleCellExperiment objects are the small dolls contained within the larger QFeatures doll.

The [`SingleCellExperiment`](http://bioconductor.org/packages/release/bioc/html/SingleCellExperiment.html) class provides a dedicated framework for single-cell data, acting as an interface to cutting-edge methods for processing, visualizing, and analyzing single-cell data.
The [`QFeatures`](https://rformassspectrometry.github.io/QFeatures/) class, on the other hand, is a data framework specifically designed to manipulate and process MS-based quantitative data. It preserves the relationship between different levels of information, such as peptide to spectrum match (PSM) data, peptide data, and protein data. Additionally, the QFeatures package offers an interface to various utility functions that streamline the processing of MS data. For further details on MS data analysis tools, refer to the [RforMassSpectrometry project](https://www.rformassspectrometry.org/).

# Before You Start

Before diving into the SCP data analysis using `scpGUI`, it's essential to note that this app serves as an interface to the `scp` package. When errors occur during the analysis, `scpGUI` will simply report the error message. In such cases, users are encouraged to refer to the documentation of the associated function within the `scp` package to identify and resolve the issue. The error messages provide valuable clues, and searching the `scp` package documentation will help users troubleshoot effectively.

# Installation and Launch
## Installation

```{r, eval = FALSE}
# Check if remotes is installed. Otherwise install it.
if (!require("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
# Install the package
remotes::install_github("UCLouvain-CBIO/scpGUI")
# Load the package
library(scpGUI)
```

## Application launch

```{r, eval = FALSE}
app <- scpGUI(example = FALSE)

if (interactive()) {
    shiny::runApp(app)
}
```

# App sections

When launching `scpGUI`, the user can see section of the app on the right sidebar. Each section will serve one step of the workflow.
Each section is documented in its own vignette:

- [Import Section](https://uclouvain-cbio.github.io/scpGUI/articles/import_section.html)

# Citation

```{r citation}
citation("scpGUI")
```

# References
